---
title: "Getting Started with isimip"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with isimip}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

The `isimip` package provides an R interface to the [ISIMIP (Inter-Sectoral Impact Model Intercomparison Project) Repository](https://data.isimip.org). This vignette demonstrates the basic usage of the package.

## Installation

```{r setup, eval = FALSE}
# Install from GitHub
devtools::install_github("limnotrack/isimip-client")

# Load the package
library(isimip)
```

## Creating a Client

The main interface is the `ISIMIPClient` S4 class:

```{r}
library(isimip)

# Create a client with default settings
client <- ISIMIPClient()
```

You can customize the API endpoints and version:

```{r}
# Use custom URLs
client <- ISIMIPClient(
  data_url = "https://data.isimip.org/api/v1",
  files_api_url = "https://files.isimip.org/api/v2",
  files_api_version = "v2"
)

# Add authentication if needed
client <- ISIMIPClient(
  auth = list(username = "your_username", password = "your_password")
)
```

## Searching for Datasets

### Basic Search

Search using a query string:

```{r}
# Search for datasets
datasets <- datasets(client, query = "gfdl-esm4 ssp370 pr")

# Check the number of results
cat("Found", datasets$count, "datasets\n")

# View first result
str(datasets$results[[1]])
```

### Search with Specifiers

Use specific parameters to narrow your search:

```{r}
datasets <- datasets(client,
  simulation_round = "ISIMIP3b",
  product = "InputData",
  climate_forcing = "gfdl-esm4",
  climate_scenario = "ssp370",
  climate_variable = "pr"
)
```

### Search by Path

Search for datasets in a specific path:

```{r}
datasets <- datasets(client,
  path = "ISIMIP3b/InputData/climate/atmosphere/bias-adjusted/global/daily/ssp370/GFDL-ESM4/"
)
```

### Pagination

By default, the client fetches all results. For manual pagination:

```{r}
# Get first page
page1 <- datasets(client,
  query = "gfdl-esm4 ssp370 pr",
  paginate = TRUE,
  page = 1,
  page_size = 10
)

# Get second page
page2 <- datasets(client,
  query = "gfdl-esm4 ssp370 pr",
  paginate = TRUE,
  page = 2,
  page_size = 10
)
```

## Searching for Files

Similar methods are available for searching files:

```{r}
# Search for files
files <- files(client,
  simulation_round = "ISIMIP3b",
  climate_forcing = "gfdl-esm4",
  climate_scenario = "ssp370",
  climate_variable = "pr"
)

# Get a specific file by ID
file_detail <- file(client, pk = "some-uuid-here")
```

## Processing Operations

The Files API allows you to perform operations on data before downloading.

### Select Operations

Extract data for specific regions or points:

```{r}
# Select data for a bounding box
result <- select_bbox(client,
  paths = c(
    "ISIMIP3b/InputData/climate/atmosphere/bias-adjusted/global/daily/ssp370/GFDL-ESM4/gfdl-esm4_r1i1p1f1_w5e5_ssp370_pr_global_daily_2015_2020.nc"
  ),
  west = -20, east = 20,
  south = -10, north = 10,
  poll = 4  # Poll every 4 seconds for job completion
)

# Select data for a point
result <- select_point(client,
  paths = c("path/to/file.nc"),
  lat = 6.25, lon = 18.17,
  poll = 4
)
```

### Mask Operations

Apply masks to limit data to specific regions:

```{r}
# Mask to a bounding box
result <- mask_bbox(client,
  paths = c("path/to/file.nc"),
  west = -20, east = 20,
  south = -10, north = 10,
  poll = 4
)

# Mask to a country (using ISO 3166-1 alpha-3 code)
result <- mask_country(client,
  paths = c("path/to/file.nc"),
  country = "bra",  # Brazil
  poll = 4
)

# Mask to land areas only
result <- mask_landonly(client,
  paths = c("path/to/file.nc"),
  poll = 4
)
```

### Cutout Operations

Extract subsets of data:

```{r}
# Cut out a rectangular region
result <- cutout_bbox(client,
  paths = c("path/to/file.nc"),
  west = 5.8, east = 10.6,
  south = 45.8, north = 47.9,
  poll = 4
)

# Cut out data at a point
result <- cutout_point(client,
  paths = c("path/to/file.nc"),
  lat = 46.5, lon = 8.0,
  poll = 4
)
```

### Custom Operations with Shapefile

You can use custom shapefiles for masking:

```{r}
result <- mask_shape(client,
  paths = c("path/to/file.nc"),
  shape = "~/data/shapes/countries.zip",
  layer = 5,  # Layer index or name
  poll = 4
)
```

### Complex Operation Chains

Chain multiple operations together:

```{r}
operations <- list(
  list(
    operation = "cutout_bbox",
    bbox = c(5.8, 10.6, 45.8, 47.9)
  ),
  list(
    operation = "mask_country",
    country = "che"  # Switzerland
  )
)

result <- submit_job(client,
  paths = c("path/to/file.nc"),
  operations = operations,
  poll = 4
)
```

## Downloading Results

Once a job is complete, download the result:

```{r}
if (!is.null(result$file_url)) {
  # Download to current directory
  download(client, result$file_url)
  
  # Download to specific directory
  download(client, result$file_url, path = "downloads")
  
  # Download with validation and extraction
  download(client,
    result$file_url,
    path = "downloads",
    validate = TRUE,  # Validate checksum
    extract = TRUE    # Extract zip files
  )
}
```

## Working with Results

The results from operations are lists that you can process in R:

```{r}
# Check job status
cat("Job ID:", result$id, "\n")
cat("Status:", result$status, "\n")

if (result$status == "finished") {
  cat("Download URL:", result$file_url, "\n")
  cat("Metadata:", result$meta, "\n")
}
```

## Legacy API (v1)

For backward compatibility, you can use the v1 API:

```{r}
# Create v1 client
client_v1 <- ISIMIPClient(
  files_api_url = "https://files.isimip.org/api/v1",
  files_api_version = "v1"
)

# Use v1 methods
result <- client_v1$select(
  paths = c("path/to/file.nc"),
  bbox = c(-10, 10, -20, 20)  # south, north, west, east
)

result <- client_v1$mask(
  paths = c("path/to/file.nc"),
  country = "bra"
)

result <- client_v1$cutout(
  paths = c("path/to/file.nc"),
  bbox = c(-10, 10, -20, 20)
)
```

## Error Handling

The package provides informative error messages:

```{r, error = TRUE}
# Trying to use v2 method with v1 client
client_v1 <- ISIMIPClient(files_api_version = "v1")
client_v1$select_bbox(
  paths = c("path/to/file.nc"),
  west = -20, east = 20, south = -10, north = 10
)
# Error: This method is only available in v2 of the Files API
```

## Next Steps

- Check out the [package website](https://limnotrack.github.io/isimip-client/) for more examples
- Read the function documentation: `?ISIMIPClient`
- Visit the [ISIMIP Data Portal](https://data.isimip.org) to explore available data
